Automated Formal Synthesis of Provably Safe Digital Controllers for Continuous Plants

### 2019-07-20 automatica/response.tex

\usepackage{amssymb,latexsym,amsfonts,amsmath, amsthm,color,mathrsfs}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{todonotes}
%\usepackage{utf8}
\usepackage{fullpage}
\def\SS#1{{\textcolor{red}{ {\bf SS:} #1}}}
\renewcommand{\b}{\textcolor{blue}}
%\newcommand{\red}{\textcolor{red}}
%
\newcommand{\vold}[1]{\colorbox{red!20}{\bfseries #1}}
\newcommand{\vnew}[1]{\colorbox{blue!30}{\bfseries #1}}

\renewcommand{\t}{\new{(t)}}
\newcommand{\tp}{\new{(t+1)}}
\newcommand{\fei}[1]{{\color{purple}#1 \color{black}}}
\newcommand{\rev}[1]{\textcolor{blue}{#1}}

\begin{document}
\begin{center}
{\Large Responses to the Reviews}\\\vspace{0.3cm}
\today
\end{center}
\vspace{1cm}
Dear Associate Editor and Reviewers,\vspace{0.3cm}
\\
We would like to deeply thank you for reviewing our paper and providing us with the constructive comments and suggestions. We have carefully addressed
all the comments and incorporated them in the revised manuscript.

We present below a detailed discussion of all comments, questions, and concerns raised by the reviewers. The major modifications are highlighted in
blue in the revised manuscript. 
\todo{keep next?} For clarity,  we also use \vold{v1} and \vnew{v2} in the response letter for referring respectively to the previous
and current versions of the manuscript.
%I do not like color coding that much. And definitely not with red and blue.
\\
\\
Best regards,\vspace{0.1cm}
\\
The Authors
%
\newpage

@@

{\bf Comment: }{\itshape This paper presents an approach for synthesizing 
digital/software controllers for continuous plants (linear time invariant / LTI) 
such that they satisfy safety/stability requirements, as would show up in closed-loop 
control of cyber-physical systems (CPS). The approach extends earlier work of the authors, 
the primary of which in my view is toward floating-point implementation as opposed to only fixed-point. 
I did not entirely buy that using CEGIS is a contribution, as from looking over the earlier related publications 
(e.g., ASE'17, CAV'17, HSCC'17, SYNT'18 abstract), all of them use CEGIS.}

\vspace{1em}

{\bf Response: }

This journal submission encompasses and summarises the mentioned conference articles (CAV, HSCC, ASE) or workshop abstract (SYNTH). 
It is a more mature, complete, and comprehensive presentation of the conference material. 
Of these conference articles, the CAV/HSCC contributions provided an early presentation of the theory underlying the approach, 
whereas the ASE was a tool paper.  
The present contribution has new or updated experimental outcomes. 
The application of the CEGIS architecture to the synthesis of digital controllers for continuous plants is indeed a novel contribution of this line of
work.

@@

Fixed this now

\todo[inline]{Elizabeth: say what the problem was, fix this if possible, then thank the reviewer for pointing this out. }

\vspace{2em}
{\bf Comment: } {\itshape On the control theory side, the approach uses standard results to synthesize stabilizing controllers for LTI systems,
specifically linear state feedback controllers (u=Kx with a possible reference signal to track), under the assumption the plant is controllable. It
seems the approach requires full-state observability as well, which is a fairly standard assumption as one can typically implement various filters /
estimators (Kalman, Leuenberger, etc.) to ensure state observability assuming detectability of the plant, but the authors should clarify this point to
make the assumptions more precise.}

\vspace{1em}
{\bf Response: }
We do need full observability of the (states of the) model, and have clarified this in Section 3.1, as follows: {\it We assume full observability of
the states of the model, namely the output states correspond with the model variables.  } We recall this assumption later, throughout the manuscript.

@@

\vspace{1em}

{\bf Response: }

We have added the completeness threshold for each benchmark in the table. Note that the completeness threshold is dependent on the controller, 
and a controller with a smaller completeness threshold may exist.

The proof of Theorem 2 has been elaborated and has been made more detailed.

@@

\vspace{2em}

{\bf Comment: }
{\itshape I would like to see some additional discussion and/or comparison to hybrid systems reachability. While the focus of this work is on
lower-level
software synthesis, if the approach is to be believed in the context of closed-loop systems satisfying specifications, then I would like some
discussion/comparison to that vast existing literature. Specifically, I presume the reason the benchmarks are not scalable beyond ~n=4 dimensionality
(as the n=5 benchmark appears to timeout), which is trivially small at this stage for linear systems (e.g., given a recent HSCC'19 paper on verifying
$R^{billion}$ linear systems, or at least n=100 to n=1000 with standard approaches like SpaceEx), is due to overapproximation error growth with the
interval representation of plant behaviors, whereas most scalable approaches utilize better data structures for representing geometric sets
(zonotopes, support functions, star sets, Taylor models, etc.). This likewise limits the possible applicability of the
approach for nonlinear plants, although there are a host of technical issues in that potential extension. }

\vspace{1em}
{\bf Response: }

We have reviewed relevant literature on controller synthesis based on finite abstractions, cf. Belta, Tabuada, Zamani, et al.. We have clarified the differences with our approach in the manuscript.

Furthermore, most literature on reachability analysis tackles a different problem, namely one of verification, rather than one of control synthesis (which this work
focuses on). Admittedly, reachability analysis tool scale quite well (as the reviewer points out), however this is under clear assumptions on model dynamics. 
Whilst algorithmically very advanced, notice that only a few of the available software tools for reachability analysis are actually numerically sound.

So one might argue that reachability analysis tools could be embedded within the CEGIS framework (incidentally, this was partly attempted in our CAV17 contribution), 
however they currently lack full soundness: namely they 
are algorithmically sound, but not numerically sound. On the contrary our work is sound on both sides.

@@

\vspace{2em}
{\bf Comment: }
{\itshape Finally, while checking the $2^n$ vertices
are sufficient (theorem 1), I doubt this is necessary, as e.g. the star set approaches avoid this blow-up that would also limit the dimensionality
the approach can work toward.}
\vspace{1em}

{\bf Response: }

It is probably true that the check on all vertices might not be necessary in general, that that this check might be simplified in special cases
or under special conditions. The check however, as it currently is implemented, does not represent the bottleneck of our overall method, so we can
safely rely on its current form, as it stands.  \todo{Add note to manuscript?}

@@

The same theorem is also present in [1]. The proof, upon careful inspection, is not a proof of the statement given in the theorem. It only proves
that the reachable set is the convex set obtained by the trajectories of the corner states. However, it does not prove that if the trajectories are
safe, their convex combination is also safe. In fact, as given in the above counterexample, the vertices of polytope are safe does not mean that the
convex combination is safe.

I think this bug needs to be fixed in the BMC part of the verifier. Instead of checking the individual trajectories, you should in fact compute the
convex combination of trajectories (reachable set) and then perform the verification.}

\vspace{1em}
{\bf Response: }
\todo[inline]{Alessandro: Response to this (Theorem 1)}
\vspace{2em}

Many thanks for this note. 
Notice that already in the previous manuscript we had focused on special shapes of initial and safe sets, namely (convex) hyper-boxes, which include the
origin. 
In the amended draft, we have better clarified these assumptions, and 
we have emphasised that the initial set is assumed to lie within the safe set: indeed, the contrary would trivially result in an unsafe condition. 
In conclusion, recall that the sets are hyper-boxes, thus convex, and that in view of the linear dynamics of the considered models, in practice both 
are assumed to contain the origin.

@@

\section{Response to Reviewer (4)}

{\bf Comment:  } {\itshape The safety property in Eqn 3 focuses on interval constraints. One can generate auxiliary variables such that range
constraints over them can encode
more complex safety properties. It would be nice to bring this up in the discussion after equation 3 that $x_i$ need not be just $i$-th component of
the
state.  Can these also incorporate temporal properties (say STL properties with bounded time-horizon) without having to unroll the transition and
create auxiliary variables that track history explicitly?}

\vspace{1em}
{\bf Response:}
 It is true that we would be able to encode more complex properties by introducing auxiliary variables; 
 however,  Theorem 2 is dependent on the safe set being convex, and so this proof would not necessarily hold with more complex properties. 
 It would be possible to construct time-bounded properties as automata that track history explicitly, but we can't see how
it would be possible to avoid unwinding the transition relation using this method. In the case of the example given by the reviewer, an STL property with bounded time-horizon, unwinding
the transition relation would not pose much of a problem anyway as a time-bounded property would only require unwinding up to the time-bound.

@@

%\todo[inline]{Eli: can we generalise according to the suggestion of the reviewer?}

\vspace{2em}
{\bf Comment: } {\itshape The paper would benefit with a restructuring of the discussion. Figure 2 and 3 will be more useful
earlier in the paper (Section 1 or 3).}

\vspace{1em}
{\bf Response: }
We have moved these figures to Section 1 and incorporated them into the Introduction. We are still referring back to these figures from Section 5.

@@

\vspace{2em}
{\bf Comment: } {\itshape 
Could you add one example to the end of Section 4 to show $R_1$, $R_2$, $R_3$, $R_4$ for a simple LTI system?

\vspace{1em}
{\bf Response: }

\todo[inline]{Eli: can you clarify? Again, this is just Jury's criteria, we don't really have space to go over examples of basic control theory }

\todo{can we at least say that the values can be obtained form the SW tool, and say where in the tool they are computed? }

### 2019-07-23 automatica/response.tex

This journal submission encompasses and summarises the mentioned conference articles (CAV, HSCC, ASE) or workshop abstract (SYNTH). 
It is a more mature, complete, and comprehensive presentation of the conference material. 
Of these conference articles, the CAV/HSCC contributions provided an early presentation of the theory underlying the approach, 
whereas the ASE was a tool paper.  
The present contribution has new or updated experimental outcomes. 
The application of the CEGIS architecture to the synthesis of digital controllers for continuous plants is indeed a novel contribution of this line of
work.

\vspace{2em}
{\bf Comment: } {\itshape The results are evaluated on about a dozen and a half examples from the controls literature all with relatively small
dimensionality (2 to 5 real variables in the continuous state space), and while the authors claim the software artifacts and benchmarks are available
for evaluation, when this reviewer browsed to the link (http://www.cprover.org/DSSynth/), a page with "TBD" was shown. I did a little digging and
eventually found what I presume is the same / similar artifacts here: https://github.com/ssvlab/dsverifier and/or http://dsverifier.org/}

\vspace{1em}
{\bf Response: }

We apologise for this error and thank the reviewer for pointing it out. The DSSynth download from http://dsverifier.org/ is indeed the same code 
as our implementation for this paper, but without the experimental results and benchmarks. The correct link is now functional: http://www.cprover.org/DSSynth/.

\todo[inline]{fix link }

@@

It is probably true that the check on all vertices might not be necessary in general, that that this check might be simplified in special cases
or under special conditions. The check however, as it currently is implemented, does not represent the bottleneck of our overall method, so we can
safely rely on its current form, as it stands.  \todo{Add note to manuscript?}

\vspace{2em}

{\bf Comment: }
{\itshape While I may have missed it, I didn't see how the controller gain K is initially selected. I presume under the given assumptions a standard
approach
such as pole placement, but this should be clarified}
\vspace{1em}
{\bf Response: }

The controller gain is initially nondeterministically selected by the bounded model checking synthesis step, i.e., the SAT solver returns the first
assignment to the controller
values it finds
that satisfies the boolean formula which encodes the specification, by using Conflict Driven Clause Learning.
We have added the following sentence to clarify this:

\rev{The bounded model checker selects the controller $K$ values using a satisfiability solver, i.e.,
by constraint solving. As such, we do not need to consider traditional control approaches such as pole placement.}

### 2019-07-23 automatica/paper.tex
temporary commit

@@

We fix an index $k$, 
and we synthesize a safe controller by
unfolding the transition system (i.e., the closed-loop model) $k$ steps and by selecting a controller $K$ and a single initial state, 
such that the states at each step do not violate the safety requirement (see Section~\ref{sec:safety}).    
That is, we ask the bounded model checker~\cite{ClarkeKL04} if there exists a $K$ that is safe for at least one $x_0$ in our set of all possible initial states, and a given fixed-point precision for the controller, and where the input signal remains within the specified bounds. 
This approach is sound, i.e., the controller produced is guaranteed to be safe, if the current $k$ is greater than the completeness threshold (see later step).  
We also assume some finite precision for the model and a given time discretisation, as described in section~\ref{sec:model}.  
We use a fixed-point precision in this description, given by $\langle I_p,F_p\rangle$, but if we are considering a floating-point controller we will instead model the plant with a floating-point precision, $\langle E_p,M_p\rangle$.
The checks that these assumptions hold are performed by the subsequent {\sc verify} stages.

>>>>

A large body of work exists on evaluating fixed-point errors and bridging the gap between real value
and fixed-point values in synthesis in other application areas and using other
algorithmic techniques; 
Genetic programming has been used to minimise the error in synthesised fixed-point programs~\cite{Darulova};
Smoothed proof search reduces the problem of boolean  a sequence of optimization problems

@@

We fix an index $k$, 
and we synthesize a safe controller by
unfolding the transition system (i.e., the closed-loop model) $k$ steps and by selecting a controller $K$ and a single initial state, 
such that the states at each step do not violate the safety requirement (see Section~\ref{sec:safety}).    
That is, we ask the bounded model checker~\cite{ClarkeKL04} if there exists a $K$ that is safe for at least 
one $x_0$ in our set of all possible initial states, and a given fixed-point precision for 
the controller, and where the input signal remains within the specified bounds. 
\rev{The bounded model checker selects the controller $K$ values using a satisfiability solver, i.e.,
by constraint solving. As such, we do not need to consider traditional control approaches such as pole placement.}
This approach is sound, i.e., the controller produced is guaranteed to be safe, if the 
current $k$ is greater than the completeness threshold (see later step).  
We also assume some finite precision for the model and a given time discretisation, as described in section~\ref{sec:model}.  
We use a fixed-point precision in this description, given by $\langle I_p,F_p\rangle$, but if we are considering a floating-point controller we will instead model the plant with a floating-point precision, $\langle E_p,M_p\rangle$.
The checks that these assumptions hold are performed by the subsequent {\sc verify} stages.
